Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2021-02-26 Nicolas.tex
2 columns. DONE!!
Started final changes for acceptance

\documentclass[referee]{aa}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties rises, notably that of SN~Ia population evolution. In this
    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
    lightcurve stretch, a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to strongly correlate with the
    SN environment, notably with stellar age tracers. We model the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted by \cite{rigault2020}, and
    assuming constant underlying stretch distribution for each age population
    consisting of Gaussian mixtures. We test our prediction against published
    samples cut to have marginal magnitude selection effects, so that any
    observed change is indeed of astrophysical and not observational in origin.
    In this first study, there are indications that the underlying SN~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a better description of the data than any time-constant
    model, including the sample-based asymmetric distributions often used to
    correct Malmquist bias \textbf{at a significance of over 5 $\sigma$}. The
    favored underlying stretch model is the bimodal one derived from
    \cite{rigault2020}, composed of a high-stretch mode shared by both young and
    old environments, and a low-stretch mode exclusive to old environments. The
    precise impact of the redshift evolution of the SN Ia population intrinsic
    properties on cosmology remains to be studied. Yet, the astrophysical drift
    of the SN stretch distribution does affect current Malmquist bias
    corrections and thereby distances derived from SN affected by observational
selection effects. We highlight that such a bias will increase with surveys
covering increasingly larger redshift ranges, which is particularly important
for LSST.}

@@

    \item Redshift-independent models are quantitatively excluded as
        suitable descriptions of the data relative to our Base model. This model
        assumes that: (1) the younger population has a unimodal Gaussian stretch
        distribution, while the older population stretch distribution is
        bimodal, one mode being the same as the young one; (2) the evolution of
        the relative fraction of younger and older SNe~Ia follows the prediction
        made in \cite{rigault2020}. This second result further supports
        the existence of both young and old SN~Ia populations, in agreement with
        rate studies \cite{mannucci2005, scannapieco2005, sullivan2006,
        aubourg2008}.

>>>>

\documentclass[]{aa}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties rises, notably that of SN~Ia population evolution. In this
    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
    lightcurve stretch, a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to strongly correlate with the
    SN environment, notably with stellar age tracers. We model the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted \textbf{in a previous study},
    and assuming constant underlying stretch distribution for each age
    population consisting of Gaussian mixtures. We test our prediction against
    published samples cut to have marginal magnitude selection effects, so that
    any observed change is indeed of astrophysical and not observational in
    origin. In this first study, there are indications that the underlying
    SN~Ia stretch distribution is evolving as a function of redshift, and that
    the young/old drifting model is a better description of the data than any
    time-constant model, including the sample-based asymmetric distributions
    often used to correct Malmquist bias at a significance of over 5 $\sigma$.
    The favored underlying stretch model is \textbf{a bimodal one}, composed of
    a high-stretch mode shared by both young and old environments, and a
    low-stretch mode exclusive to old environments. The precise impact of the
    redshift evolution of the SN Ia population intrinsic properties on cosmology
    remains to be studied. Yet, the astrophysical drift of the SN stretch
    distribution does affect current Malmquist bias corrections and thereby
    distances derived from SN affected by observational selection effects. We
highlight that such a bias will increase with surveys covering increasingly
larger redshift ranges, which is particularly important for LSST.}

@@

    \item Redshift-independent models are quantitatively excluded as
        suitable descriptions of the data relative to our Base model. This model
        assumes that: (1) the younger population has a unimodal Gaussian stretch
        distribution, while the older population stretch distribution is
        bimodal, one mode being the same as the young one; (2) the evolution of
        the relative fraction of younger and older SNe~Ia follows the prediction
        made in \cite{rigault2020}. This second result further supports
        the existence of both young and old SN~Ia populations, in agreement with
        rate studies \citep{mannucci2005, scannapieco2005, sullivan2006,
        aubourg2008}.
